---
title: "Data preparation"
author: "Camille Crapart"
date: "25 2 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = T, collapse = T,  message = F, warning = F, error = F)
```

```{r libraries}

library(readxl)
library(dplyr)
library(ggplot2)
library(dlookr)

library(sf)
library(raster)
library(spatialreg)
library(spdep)

library(MASS)

library(factoextra)

memory.limit(size = 1e+10)

id.names <- c("station_id","station_code","nve_vatn_nr","nve_vassdrag_nr", "CBA_Lake_ID", "STATION_ID", "statn_d","longitude","latitude", "region")
is_constant <- function(x) length(unique(x)) == 1L 
```

# Data preparation 

## Import data from NIVA

```{r load-data-niva, eval = F}
id.niva <- read_xlsx("id_niva.xlsx")

catchment.poly <- st_read("catchments_poly/lakes1000cat.shp")
niva.stations <- st_read("catchments_poly/lakes1000stations.shp")
niva.data <- read_xlsx("niva_selection.xlsx", sheet = "niva")

all.niva <- st_drop_geometry(niva.stations) %>% merge(niva.data, by.x = "sttn_cd", by.y = "station_code") %>% merge(catchment.poly, by.x = "statn_d", by.y= "station_id")

#bdg.stations <- id.niva %>% merge(niva.stations, by.x = "STATION_ID", by.y = "sttn_cd") %>% merge(niva.data, by.x = "STATION_ID", by.y = "station_code")
bdg.niva <- id.niva %>% merge(all.niva, by.x = "STATION_ID", by.y = "sttn_cd") %>% st_as_sf()

missing.bdg <- id.niva %>% filter(!STATION_ID %in% bdg.niva$STATION_ID)

saveRDS(bdg.niva, "bdg.niva.rds")
```

```{r plot-niva-catchments}
ggplot(bdg.niva)+geom_sf(aes(fill = as.factor(region), col = as.factor(region)))+
  scale_color_brewer(palette = "Spectral", aesthetics = c("col","fill"))+
  theme_minimal(base_size = 8)

ggplot(bdg.niva)+geom_point(aes(x = longitude, y = latitude, col = as.factor(region)))+
  scale_color_brewer(palette = "Spectral", aesthetics = c("col","fill"))+
  theme_minimal(base_size = 8)

```

Region 1: 
Hedmark
Oppland

Region 2: 
Telemark
Buskerud
Vestfold
Østfold
Akershus
Hedmark
Oppland

Region 3
Telemark
Oppland
Buskerud
Sogn of Fjordane
Hordaland

Region 4
Aust-Agder
Telemark
Vest-Agder

Region 5
Rogaland
Vesst-Agder
Hordaland

Region 6
Hordaland
Rogaland
Sogn og Fjordane   

Region 7
Hordaland
Sogn og Fjordane

Region 8 
Sør-Trøndelag
Møre og Romsdal
Nord-Trondelag

Region 9 
Troms
Nordland
Finmark

Region 10
Finmark

Missing data comes from the excel file from Heleen.

## Extract NDVI 2019 

```{r extract-ndvi, eval = F}
copernicus.files <- list.files(path = "Summer_NDVI_2019", pattern = "1.nc", full.names = T)

copernicus.stack <- raster::stack(copernicus.files, varname = "NDVI")
copernicus.fennoscandia <- raster::crop(copernicus.stack, c(0,35,55,73))
copernicus.mean <- copernicus.fennoscandia %>% stackApply(indices = c(1:length(copernicus.files)), fun = mean, na.rm = T)

copernicus.fen <- copernicus.fennoscandia %>% mean()

niva.ndvi <- raster::extract(copernicus.fen, catchments.bdg, sp = T, fun = mean, df = T, na.rm = T)
names(niva.ndvi) <- c("station_id", "ndvi")
saveRDS(niva.ndvi, "Summer_NDVI_2019/niva.ndvi.rds")
```

## Soil

Soil thickness: extracted from ORNL DAAC: https://daac.ornl.gov/SOILS/guides/Global_Soil_Regolith_Sediment.html 

Chose the average value: "average_soil_and_sedimentary-deposit_thickness.tif"
-> A grid that averages soil and sedimentary deposit thicknesses in meters for users who want a single thickness value that averages across upland hillslope and valley bottoms.

```{r extract-soil-depth, eval = F}
soil <- raster::raster("DAAC/average_soil_and_sedimentary-deposit_thickness.tif")
catchment.poly <- st_read("catchments_poly/lakes1000cat.shp")
soil.df <- extract(soil, catchment.poly, sp = T, df = T, na.rm = T, fun = mean)
names(soil.df) <- c("station_id","soil")
saveRDS(soil.df, "soil.df.rds")
```

## Gather dataset

```{r niva-sel, eval = F}
bdg.niva <- readRDS("bdg.niva.rds")
explore.niva <- bdg.niva[,-which(names(bdg.niva) %in% id.names)] %>% select_if(is.numeric) %>% st_drop_geometry()
explore.niva <- explore.niva[,apply(explore.niva, 2, var, na.rm = T) != 0]

niva.prcomp <- prcomp(~., data = explore.niva, center = T, scale. = T)
niva.pca.var <- fviz_pca_var(niva.prcomp, col.ind = "contrib")
print(niva.pca.var)

apply(explore.niva, 2, find_na)
```

```{r create-dataset, eval = F}
niva.ndvi <- readRDS("Summer_NDVI_2019/niva.ndvi.rds")
bdg.niva <- readRDS("bdg.niva.rds")
soil <- readRDS("soil.df.rds")
nor <- st_read("Country_shapefile/norway.shp") %>% st_transform(st_crs(bdg.niva))
  
rr <- read_xlsx("RR1000s.xlsx")
rr$...1 <- NULL

keep <- c("latitude", "longitude", "region", "ph_2019","tg_6m_19","kond_2019","toc_2019","cndom_corr_2019", "era_prec_2019", "rr_2m_19","mean_elev","nilu_ndep_2012_2016","nilu_sdep_2012_2016","freshwater","forest","agriculture","peat","basin_area_km2")

bdg.data <- bdg.niva[,which(names(bdg.niva) %in% c(id.names,keep))] %>% merge(dplyr::select(rr,c("Sample_ID","dmax","hwidth")), by.x = "CBA_Lake_ID", by.y = "Sample_ID") %>%  merge(niva.ndvi, by.x = "statn_d", by.y = "station_id") %>% merge(soil, by.x = "statn_d", by.y = "station_id")

bdg.data$bdg <- bdg.data$dmax/bdg.data$toc_2019

saveRDS(bdg.data,"bdg.data.rds")
```


# Analysis

# Bdg by region

```{r boxplot-by-region, fig.dim = c(20,10)}
bdg.data <- readRDS("bdg.data.rds")

ggplot(bdg.data)+geom_boxplot(aes(y=bdg))+facet_grid(cols=vars(region))+theme_minimal()
ggplot(bdg.data)+geom_boxplot(aes(y=hwidth))+facet_grid(cols=vars(region))+theme_minimal()
ggplot(bdg.data)+geom_boxplot(aes(y=dmax))+facet_grid(cols=vars(region))+theme_minimal()
```

## PCA

The PCA shows that the different variables for runoff, temperature and atmospheric deposition point to the same directions. Only the parameter with the least NA from each category is kept.

* rr_2m_19 for runoff
* tg_6m_19 (most correlated with laketemp_19)
* nilu_ndep_2012_2016 also for sdep

The data from 1995 are not useful and include many NA.

"Dyp" (deepness) also have many missing data (18). 

We keep toc and cndom for now. 

```{r check-selection}
bdg.data <- readRDS("bdg.data.rds")

keep <- c("bdg","dmax","hwidth","ph_2019","tg_6m_19","kond_2019","toc_2019","cndom_corr_2019", "era_prec_2019", "rr_2m_19","mean_elev","nilu_ndep_2012_2016","nilu_sdep_2012_2016","freshwater","forest","agriculture","peat","basin_area_km2")

sel2 <- bdg.data[,keep] %>% st_drop_geometry()
pca2 <- prcomp(~., data = sel2, center = T, scale. = T)
plot2 <- fviz_pca_var(pca2, col.ind = "contrib")
print(plot2)
```

## Correlation

```{r correlations}
bdg.data <- readRDS("bdg.data.rds")

matrix <- bdg.data[,-which(names(bdg.data) %in% id.names)] %>% st_drop_geometry()

cor.pred <- cor(matrix, use = "pairwise.complete.obs")

png("pred.corrplot.png")
corrplot::corrplot(cor.pred, type = "lower", tl.cex = 0.8)
dev.off()

corrplot::corrplot(cor.pred, type = "lower", tl.cex = 0.8)
```

## Transform dataset

```{r skewness, eval = F}
bdg.data <- readRDS("bdg.data.rds")

bdg.df <- bdg.data[,- which(names(bdg.data) %in% id.names[-which(id.names %in% c("CBA_Lake_ID","longitude","latitude","region"))])] %>% st_drop_geometry()

skewed.data <- find_skewness(bdg.df, index = F)

bdg.trans <- bdg.df
for(i in skewed.data){
  newname <- paste("log",i, sep = ".")
  oldnames <- names(bdg.trans)
  bdg.trans <<- cbind(bdg.trans, log(bdg.trans[[i]])) %>% setNames(c(oldnames,newname))
}

# bdg.trans <- bdg.data %>% mutate_at(.funs = transform, method = "log", .vars = skewed.data) %>% st_drop_geometry()
# sapply(bdg.trans,"class")
# bdg.trans[,which(names(bdg.trans) %in% skewed.data)] <- as.numeric(bdg.trans[,which(names(bdg.trans) %in% skewed.data)])
# 
# skewed.index <- which(names(bdg.trans) %in% skewed.data)
# names(bdg.trans)[skewed.index] <- paste("log",skewed.data,sep = ".")

find_na(bdg.trans, index = F)
bdg.trans <- bdg.trans %>%  filter(is.na(log.nilu_ndep_2012_2016) == F) %>% filter(is.na(log.cndom_corr_2019)==F) %>% filter(log.soil != -Inf) %>% filter(log.hwidth != -Inf)

bdg.trans$log.agriculture <- NULL
bdg.trans$log.forest <- NULL
bdg.trans$log.peat <- NULL
bdg.trans$log.ndvi <- NULL

saveRDS(bdg.trans,"bdg.trans.rds")
```

```{r outliers}
library(outliers)
outliers <- bdg.trans %>% find_outliers(index = F)
gt.outliers <- apply(bdg.trans[outliers],2, grubbs.test, two.sided = T)
```

## LogBdg by region

```{r boxplot-by-region, fig.dim = c(20,10)}
bdg.trans <- readRDS("bdg.trans.rds")

ggplot(bdg.trans)+geom_boxplot(aes(y=log.bdg))+facet_grid(cols=vars(region))+theme_minimal()
ggplot(bdg.trans)+geom_boxplot(aes(y=log.hwidth))+facet_grid(cols=vars(region))+theme_minimal()
ggplot(bdg.trans)+geom_boxplot(aes(y=log.dmax))+facet_grid(cols=vars(region))+theme_minimal()
```

## Model by region

```{r region-1}

bdg1 <- filter(bdg.trans,region == 1)

lm(formula = log.dmax ~ ndvi + log.nilu_ndep_2012_2016 + SUVA, data = bdg1) %>% summary()

library(MASS)
preds <- c("log.bdg","agriculture","forest","peat","log.nilu_ndep_2012_2016","ph_2019","log.kond_2019", "log.toc_2019", "log.cndom_corr_2019","log.ndvi")
full.model <- lm(log.bdg~., data = bdg1[preds]) 
summary(full.model)
step.model <- stepAIC(full.model, direction = "both")
summary(step.model)

final.model <- lm(log.bdg~mean_elev+log.basin_area_km2+log.nilu_sdep_2012_2016, data = bdg1)
summary(final.model)

```


# Extra code

## Boxplot

```{r other-boxplots}
toplot <- names(bdg.trans)
predictors.boxplots <- lapply(toplot, function(x) ggplot(bdg.trans) + geom_boxplot(aes(y = .data[[x]]), outlier.size = 0.3, lwd = 0.2) + theme_minimal(base_size = 10))

all.boxplots <- cowplot::plot_grid(plotlist = predictors.boxplots, ncol = 5)
cowplot::save_plot("predictors.boxplots.png", all.boxplots, base_asp = 0.8)
```

Outliers detection : https://statsandr.com/blog/outliers-detection-in-r/

## Paired scatterplots

```{r scatterplots}
toplot <- c("log.bdg", "log.dmax", "log.hwidth","mean_elev","agriculture","forest","peat","freshwater","log.basin_area_km2","log.soil","log.nilu_ndep_2012_2016","log.nilu_sdep_2012_2016","log.rr_2m_19","ph_2019","log.kond_2019", "log.toc_2019", "log.cndom_corr_2019","log.ndvi")
pairs(bdg.trans[toplot], cor = T, cex = 0.01)
```

```{r maps-predictors}
bdg.trans <- readRDS("bdg.trans.rds")
nor <- st_read("Country_shapefile/norway.shp")# %>% st_transform(st_crs(bdg.trans))

for(i in c("log.toc_2019","log.bdg","log.dmax","log.hwidth")){
  g <- ggplot()+geom_sf(data = nor, fill = "white")+geom_point(data = bdg.trans, aes(x = longitude, y = latitude, col = .data[[i]]))+
  scale_color_distiller(type = "seq", palette = 4, direction = 1)+theme_void(base_size = 25)+theme(legend.position = "bottom")
  ggsave(plot = g, filename = paste("map", i, "png",sep = "."))
}  
```



# Test models

```{r stepwise-regression-bdg}
library(MASS)
preds <- c("log.bdg","mean_elev","agriculture","forest","peat","freshwater","log.basin_area_km2","log.soil","log.nilu_ndep_2012_2016","log.nilu_sdep_2012_2016","log.rr_2m_19","ph_2019","log.kond_2019", "log.toc_2019", "log.cndom_corr_2019","log.ndvi")
full.model <- lm(log.bdg~., data = bdg.trans[preds])
step.model <- stepAIC(full.model, direction = "both")
summary(step.model)

final.model <- lm(log.bdg~mean_elev+log.basin_area_km2+log.nilu_sdep_2012_2016, data = bdg.trans)
summary(final.model)
final.model.toc <- lm(log.toc_2019~mean_elev+log.basin_area_km2+log.nilu_sdep_2012_2016, data = bdg.trans)
summary(final.model.toc)
```

```{r stepwise-regression-dmax}
library(MASS)
preds <- c("log.dmax","mean_elev","agriculture","forest","peat","freshwater","log.basin_area_km2","log.soil","log.nilu_ndep_2012_2016","log.nilu_sdep_2012_2016","log.rr_2m_19","ph_2019","log.kond_2019", "log.toc_2019", "log.cndom_corr_2019","log.ndvi")
full.model <- lm(log.dmax~., data = bdg.trans[preds])
step.model <- stepAIC(full.model, direction = "both")
summary(step.model)

final.model <- lm(log.dmax~mean_elev+log.nilu_sdep_2012_2016+ph_2019, data = bdg.trans)
summary(final.model)
final.model.toc <- lm(log.toc_2019~mean_elev+log.basin_area_km2+log.nilu_sdep_2012_2016, data = bdg.trans)
summary(final.model.toc)
```

```{r stepwise-regression-hwidth}
library(MASS)
preds <- c("log.hwidth","mean_elev","agriculture","forest","peat","freshwater","log.basin_area_km2","log.soil","log.nilu_ndep_2012_2016","log.nilu_sdep_2012_2016","log.rr_2m_19","ph_2019","log.kond_2019", "log.toc_2019", "log.cndom_corr_2019","log.ndvi")
full.model <- lm(log.hwidth~., data = bdg.trans[preds])
step.model <- stepAIC(full.model, direction = "both")
summary(step.model)

final.model <- lm(log.hwidth~mean_elev+forest+log.nilu_sdep_2012_2016+ph_2019, data = bdg.trans)
summary(final.model)

library(mgcv)
poly.model <- gam(log.hwidth~s(mean_elev)+s(forest)+s(log.nilu_sdep_2012_2016)+s(ph_2019), data = bdg.trans)
summary(poly.model)

```

```{r early-exit}
knitr::knit_exit(fully = F)
```

```{r lm-test}
test1 <- lm(formula = log.dmax~log.nilu_sdep_2012_2016+log.toc_2019+ph_2019, data = bdg.trans)
summary(test1)
plot(test1)

test2 <- lm(formula = log.dmax~forest+log.rr_2m_19+peat, data = bdg.trans)
summary(test2)
plot(test2)

test3 <- lm(formula = log.hwidth~era_prec_2019+log.toc_2019, data = bdg.trans)
summary(test3)
plot(test3)

test4 <- lm(formula = log.hwidth~forest+log.rr_2m_19+peat, data = bdg.trans)
summary(test4)
plot(test4)

test5 <- lm(formula = log.hwidth~log.ndvi+log.rr_2m_19+peat, data = bdg.trans)
summary(test5)
plot(test5)

test6 <- lm(formula = log.bdg~log.ndvi+log.rr_2m_19+peat+log.soil, data = bdg.trans)
summary(test6)
plot(test6)
plot(bdg.trans$log.bdg, test6$fitted.values)+abline(a=0,b=1)

```

```{r kmat}
bdg.trans.sf <- merge(bdg.data,bdg.trans, by = "CBA_Lake_ID")
bdg.centroid <- st_centroid(bdg.trans.sf, of_largest_polygon = T)

bdg.kmat <- knearneigh(bdg.centroid, k = 50) %>% knn2nb() %>% nb2listw() 
saveRDS(bdg.kmat,"bdg.kmat.rds")

```

```{r selm1}
bdg.kmat <- readRDS("bdg.kmat.rds")
selm1 <- errorsarlm(formula = log.dmax~log.nilu_sdep_2012_2016+log.toc_2019+ph_2019, data = bdg.trans, listw = bdg.kmat)

boxplot(selm1$residuals)
qqnorm(selm1$residuals)

plot(selm1$fitted.values, selm1$residuals)

comp1 <- cbind(bdg.trans$log.dmax, selm1$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp1)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```

```{r selm2}
selm2 <- errorsarlm(formula = log.dmax~log.rr_2m_19+log.ndvi+peat, data = bdg.trans, listw = bdg.kmat)

boxplot(selm2$residuals)
qqnorm(selm2$residuals)

plot(selm2$fitted.values, selm2$residuals)

comp2 <- cbind(bdg.trans$log.dmax, selm2$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp2)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) +theme_minimal()

```

```{r selm3}
selm3 <- errorsarlm(formula = log.hwidth~era_prec_2019+log.toc_2019, data = bdg.trans, listw = bdg.kmat)

boxplot(selm3$residuals)
qqnorm(selm3$residuals)

plot(selm3$fitted.values, selm3$residuals)

comp3 <- cbind(bdg.trans$log.dmax, selm3$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp3)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```
```{r selm4}
selm4 <- errorsarlm(formula = log.hwidth~forest+log.rr_2m_19+peat, data = bdg.trans, listw = bdg.kmat)

boxplot(selm4$residuals)
qqnorm(selm4$residuals)

plot(selm4$fitted.values, selm4$residuals)

comp4 <- cbind(bdg.trans$log.dmax, selm4$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp4)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```

```{r selm5}
bdg.kmat <- readRDS("bdg.kmat.rds")
selm5 <- errorsarlm(formula = log.bdg~log.ndvi+log.rr_2m_19+peat+log.soil, data = bdg.trans, listw = bdg.kmat)

boxplot(selm5$residuals)
qqnorm(selm5$residuals)

plot(selm5$fitted.values, selm5$residuals)

comp5 <- cbind(bdg.trans$log.dmax, selm5$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp5)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```


```{r skewness2}
bdg.data <- readRDS("bdg.data.rds")

bdg.df <- bdg.data[,- which(names(bdg.data) %in% id.names)] %>% st_drop_geometry()

skewed.data <- find_skewness(bdg.df, index = F)

bdg.trans2 <- bdg.data %>% mutate_at(.funs = "log", .vars = skewed.data[-which(skewed.data %in% c("agriculture","forest","peat"))]) %>% st_drop_geometry()

skewed.index <- which(names(bdg.trans2) %in% skewed.data[-which(skewed.data %in% c("agriculture","forest","peat"))])
names(bdg.trans2)[skewed.index] <- paste("log",skewed.data[-which(skewed.data %in% c("agriculture","forest","peat"))],sep = ".")

find_na(bdg.trans2, index = F)
bdg.trans2 <- bdg.trans2 %>%  filter(is.na(log.nilu_ndep_2012_2016) == F) %>% filter(is.na(log.nilu_sdep_2012_2016) == F) %>% filter(is.na(log.cndom_corr_2019)==F) %>% filter(log.hwidth != "-Inf")

```


```{r boxplot}
toplot <- names(bdg.trans2)[-which(names(bdg.trans2) %in% id.names)]
predictors.boxplots2 <- lapply(toplot, function(x) ggplot(bdg.trans2) + geom_boxplot(aes(y = .data[[x]]), outlier.size = 0.4, lwd = 0.2) + theme_minimal(base_size = 10))

all.boxplots2 <- cowplot::plot_grid(plotlist = predictors.boxplots2, ncol = 3)
cowplot::save_plot("predictors.boxplots2.png", all.boxplots2)

cor.pred <- cor(bdg.trans2[,toplot], use = "pairwise.complete.obs")
corrplot::corrplot(cor.pred, type = "lower", tl.cex = 0.8)
```



```{r outliers2, eval = F}
library(outliers)
outliers2 <- bdg.trans2[,- which(names(bdg.trans2) %in% id.names)] %>% find_outliers(index = F) 
apply(bdg.trans2[,- which(names(bdg.trans2) %in% id.names)], 2, grubbs.test, two.sided = T)

library(EnvStats)
rt.dmax <- rosnerTest(bdg.trans2$log.dmax, k = 1)
```

```{r kmat2}
bdg.trans.sf <- merge(bdg.data,bdg.trans2, by = "CBA_Lake_ID")
bdg.centroid <- st_centroid(bdg.trans.sf, of_largest_polygon = T)

bdg.neigh.set <- knearneigh(bdg.centroid, k = 50)
bdg.neigh.nb <- knn2nb(bdg.neigh.set)
bdg.kmat2 <- nb2listw(bdg.neigh.nb)
saveRDS(bdg.kmat2,"bdg.kmat2.rds")

```
```{r map-data-trans2, eval = F}
bdg.kmat2 <- readRDS("bdg.kmat2.rds")

lapply(names(select_if(bdg.trans2, is.numeric)), function (x) ggplot(bdg.trans2)+geom_point(aes(x = longitude, y = latitude, col = .data[[x]])) + scale_color_viridis_c() + theme_minimal()+borders(regions = "Norway",xlim=c(0,35),ylim=c(55,73)))
ggplot(bdg.trans2)+geom_point(aes(x = longitude, y = latitude, col = log.dmax)) + scale_color_viridis_c() + theme_minimal()

#moran.test
morani <- lapply(names(select_if(bdg.trans2, is.numeric)), function (x) moran.test(bdg.trans2[,x], listw = bdg.kmat2))
names(morani) <- names(select_if(bdg.trans2, is.numeric))
```

```{r selm1-bis}
selm1 <- errorsarlm(formula = log.dmax~log.nilu_sdep_2012_2016+log.toc_2019+ph_2019, data = bdg.trans2, listw = bdg.kmat2)

boxplot(selm1$residuals)
qqnorm(selm1$residuals)

plot(selm1$fitted.values, selm1$residuals)

comp1 <- cbind(bdg.trans2$log.dmax, selm1$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp1)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```
```{r selm2-bis, eval = F}
selm2 <- errorsarlm(formula = log.dmax~log.rr_2m_19+log.ndvi+peat, data = bdg.trans2, listw = bdg.kmat)

boxplot(selm2$residuals)
qqnorm(selm2$residuals)

plot(selm2$fitted.values, selm2$residuals)

comp2 <- cbind(bdg.trans2$log.dmax, selm2$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp2)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) +theme_minimal()

```
```{r selm-12, eval = F}
selm12 <- errorsarlm(formula = log.dmax~log.rr_2m_19+log.ndvi+peat+log.nilu_sdep_2012_2016+log.toc_2019+ph_2019, data = bdg.trans2, listw = bdg.kmat2)

boxplot(selm12$residuals)
qqnorm(selm12$residuals)

plot(selm12$fitted.values, selm12$residuals)

comp12 <- cbind(bdg.trans2$log.dmax, selm12$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp12)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) +theme_minimal()
```

```{r selm3-bis, eval = F}
selm3 <- errorsarlm(formula = log.hwidth~log.toc_2019 + ph_2019+log.cndom_corr_2019, data = bdg.trans2, listw = bdg.kmat)

boxplot(selm3$residuals)
qqnorm(selm3$residuals)

plot(selm3$fitted.values, selm3$residuals)

comp3 <- cbind(bdg.trans2$log.dmax, selm3$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp3)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```
```{r selm4-bis, eval = F}
selm4 <- errorsarlm(formula = log.hwidth~forest+log.rr_2m_19+peat, data = bdg.trans2, listw = bdg.kmat)

boxplot(selm4$residuals)
qqnorm(selm4$residuals)

plot(selm4$fitted.values, selm4$residuals)

comp4 <- cbind(bdg.trans$log.dmax, selm4$fitted.values) %>% as.data.frame() %>% setNames(c("observed", "fitted"))
ggplot(comp4)+geom_point(aes(x = observed, y = fitted))+geom_abline(slope = 1, intercept = 0) + theme_minimal()

```