---
title: "Paper3"
author: "Camille Crapart"
date: '2022-09-02'
output: html_document

bibliography: C:\\Users\\raine\\Documents\\UiO\\Bibtex\\Paper3.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F, collapse = T,  message = F, warning = F, error = F)
```

```{r libraries}

library(readxl)
library(dplyr)
library(ggplot2)
library(dlookr)

library(sf)
library(raster)
library(spatialreg)
library(spdep)

library(MASS)

library(mgcv)
library(car)

library(factoextra)
library(kableExtra)

library(corrplot)
library(cowplot)

memory.limit(size = 1e+10)

id.names <- c("station_id","station_code","nve_vatn_nr","nve_vassdrag_nr", "CBA_Lake_ID", "STATION_ID", "statn_d","longitude","latitude", "region", "region2")
is_constant <- function(x) length(unique(x)) == 1L 
```
# Intro

Aim of the study: 

* investigate links between biodegredability and catchment characteristics
* validate the biodegradability experiment on a larger scale?

Reminder: biodegradability indicators are derived from the incubation experiment conducted in 2019, with the 1000-lakes-survey samples (500 samples analysed). The experiment consists in incubating filtrated samples with addition of nutrients during 30h at 25°C.

* RR (respiration rate) is the maximum oxygen consumption rate during the incubation, in uM/h
* Bdgt (biodegration period) is the time during which oxygen consumption occurs (during exponential bacterial growth) (in h)


# Data preparation 

## Import data from NIVA

```{r load-data-niva, eval = F}
id.niva <- read_xlsx("id_niva.xlsx")

catchment.poly <- st_read("catchments_poly/lakes1000cat.shp")
niva.stations <- st_read("catchments_poly/lakes1000stations.shp")
niva.data <- read_xlsx("niva_selection.xlsx", sheet = "niva")

all.niva <- st_drop_geometry(niva.stations) %>% merge(niva.data, by.x = "sttn_cd", by.y = "station_code") %>% merge(catchment.poly, by.x = "statn_d", by.y= "station_id")

#bdg.stations <- id.niva %>% merge(niva.stations, by.x = "STATION_ID", by.y = "sttn_cd") %>% merge(niva.data, by.x = "STATION_ID", by.y = "station_code")
bdg.niva <- id.niva %>% merge(all.niva, by.x = "STATION_ID", by.y = "sttn_cd") %>% st_as_sf()

missing.bdg <- id.niva %>% filter(!STATION_ID %in% bdg.niva$STATION_ID)

saveRDS(bdg.niva, "bdg.niva.rds")
```

The lake sampled by NIVA are spread into 10 different regions, that do not correspond to the Norwegian fylke: 

* Region 1: Hedmark, Oppland
* Region 2: Telemark, Buskerud, Vestfold, Østfold, Akershus, Hedmark, Oppland
* Region 3: Telemark, Oppland, Buskerud, Sogn of Fjordane, Hordaland
* Region 4: Aust-Agder, Telemark, Vest-Agder
* Region 5: Rogaland, Vest-Agder, Hordaland
* Region 6: Hordaland, Rogaland, Sogn og Fjordane   
* Region 7: Hordaland, Sogn og Fjordane
* Region 8 : Sør-Trøndelag, Møre og Romsdal, Nord-Trondelag
* Region 9 : Troms, Nordland, Finmark
* Region 10: Finmark


The 10 regions are groupes in eco-regions. ref: @Garmo2021, appendix 1

" 2.6  Eco-regions
The ThousandLakes and TrendLakes are positioned along strong climatic, ecosystem and environmental gradients with regional contrasts in intensity of environmental pressures. Ten ecoregions have been defined that include a gradient of deposition, climate, and biogeographical conditions (Dillon et al., 2003) (Table 1), which were reduced to six in the current paper, as shown in Figure 1, i.e., South (region 4+5), East (1+2), Inland (3), West (region 6+7), Central (region 8) and North (region 9+10). The regional contrasts in lake chemistry (Figure 2, Table 2, Figure SI 2) can be partially explained by variations in air pollution, acid sensitivity, proximity to the coast, elevation, vegetation and soil characteristics (Henriksen et al., 1998). South Norway has historically received the highest loadings of atmospheric pollution, followed by east and inland Norway while central and northern Norway were much less exposed being further away from the sources (Figure 1, Table 1). The most acid-sensitive areas are also located in southern Norway (Austnes et al., 2018) and thus, the highest concentration levels of acidification-related parameters such as labile Al and SO4 and the lowest pH, alkalinity and ANC are found here. However, sources of SO4 other than air pollution such as sea-salt, marine sediments (Maxe, 2001), and geological sources also contribute to elevated concentrations of SO4, especially in Northern Norway (Wilander, 1994). Similarly, the regional contrasts in Cl-levels can be related to presence of marine clays and sea-salt aerosols, explaining why especially lakes in inland Norway, located above the marine limit, are low in Cl. Sea-salt-corrected SO4,  indicates very similar relative change compared with uncorrected SO4 (Table 2, Table SI 4)."

```{r plot-niva-catchments}
bdg.niva <- readRDS("bdg.niva.rds")

ggplot(bdg.niva)+geom_sf(aes(fill = as.factor(region2), col = as.factor(region2)))+
  scale_color_brewer(palette = "Spectral", aesthetics = c("col","fill"), name = "Region" )+
  theme_minimal(base_size = 24)

ggplot(bdg.niva)+geom_point(aes(x = longitude, y = latitude, col = as.factor(region2)))+
  scale_color_brewer(palette = "Spectral", aesthetics = c("col","fill"), name = "Region")+
  theme_minimal(base_size = 24)

```

## Extract NDVI 2019 

```{r extract-ndvi, eval = F}
copernicus.files <- list.files(path = "Summer_NDVI_2019", pattern = "1.nc", full.names = T)

copernicus.stack <- raster::stack(copernicus.files, varname = "NDVI")
copernicus.fennoscandia <- raster::crop(copernicus.stack, c(0,35,55,73))
copernicus.mean <- copernicus.fennoscandia %>% stackApply(indices = c(1:length(copernicus.files)), fun = mean, na.rm = T)

copernicus.fen <- copernicus.fennoscandia %>% mean()

niva.ndvi <- raster::extract(copernicus.fen, catchments.bdg, sp = T, fun = mean, df = T, na.rm = T)
names(niva.ndvi) <- c("station_id", "ndvi")
saveRDS(niva.ndvi, "Summer_NDVI_2019/niva.ndvi.rds")
```


## Soil

Soil thickness: extracted from ORNL DAAC: https://daac.ornl.gov/SOILS/guides/Global_Soil_Regolith_Sediment.html 

Chose the average value: "average_soil_and_sedimentary-deposit_thickness.tif"
-> A grid that averages soil and sedimentary deposit thicknesses in meters for users who want a single thickness value that averages across upland hillslope and valley bottoms.

```{r extract-soil-depth, eval = F}
soil <- raster::raster("DAAC/average_soil_and_sedimentary-deposit_thickness.tif")
catchment.poly <- st_read("catchments_poly/lakes1000cat.shp")
soil.df <- extract(soil, catchment.poly, sp = T, df = T, na.rm = T, fun = mean)
names(soil.df) <- c("station_id","soil")
saveRDS(soil.df, "soil.df.rds")
```

## Create dataset

```{r create-dataset, eval = F}
niva.ndvi <- readRDS("Summer_NDVI_2019/niva.ndvi.rds")
bdg.niva <- readRDS("bdg.niva.rds")
soil <- readRDS("soil.df.rds")
nor <- st_read("Country_shapefile/norway.shp") %>% st_transform(st_crs(bdg.niva))
  
rr <- read_xlsx("RR1000s.xlsx")
rr$...1 <- NULL

keep <- c("ph_2019","tg_6m_19","kond_2019","toc_2019","cndom_corr_2019", "era_prec_2019", "rr_2m_19","mean_elev","nilu_ndep_2012_2016","nilu_sdep_2012_2016","freshwater","forest","agriculture","peat","basin_area_km2", "tn_2019","tp_2019", "fe_2019")

bdg.data <- bdg.niva[,which(names(bdg.niva) %in% c(id.names,keep))] %>% merge(dplyr::select(rr,c("Sample_ID","dmax","hwidth")), by.x = "CBA_Lake_ID", by.y = "Sample_ID") %>%  merge(niva.ndvi, by.x = "statn_d", by.y = "station_id") %>% merge(soil, by.x = "statn_d", by.y = "station_id")

bdg.data$rrn <- bdg.data$dmax/bdg.data$toc_2019

names(bdg.data) <- c("statn_d","CBA_Lake_ID","STATION_ID","station_id","latitude","longitude","nve_vatn_nr", "nve_vassdrag_nr","region","basin_area_km2","mean_elev","agriculture","forest","peat","freshwater","precip", "ndep","sdep","runoff","temp","cond","toc","cndom","ph","tn","tp","fe","region2","rr","bdgt","ndvi","soil","geometry","rrn") 

saveRDS(bdg.data,"bdg.data.rds")
```
# Comparison 73 lakes / 500 lakes

```{r read-rr-data, fig.dim = c(18,6)}
rr100 <- read_xlsx("waterchem.xlsx") %>% dplyr::select(c("Sample_ID", "Vmax", "BdgT","TOC")) %>% setNames(c("Sample_ID","rr_100","bdgt_100","TOC_100"))
rr1000 <- readRDS("bdg.data.rds") %>% dplyr::select(c("CBA_Lake_ID", "rr", "bdgt","toc")) %>% setNames(c("Sample_ID","rr_1000","bdgt_1000", "TOC_1000"))

rr.all <- merge(rr100, rr1000, by = "Sample_ID")

rr.all$log.rr_100 <- log(rr.all$rr_100)
rr.all$log.rr_1000 <- log(rr.all$rr_1000)
rr.all$log.bdgt_100 <- log(rr.all$bdgt_100)
rr.all$log.bdgt_1000 <- log(rr.all$bdgt_1000)

p1 <- ggplot(data = rr.all) + geom_point(aes(x=rr_100, y = rr_1000))+
  labs(x="RR - NIVA 100 lakes", y = "RR - NIVA 1000 lakes")+
  theme_minimal(base_size = 25)
p2 <- ggplot(data = rr.all) + geom_point(aes(x=bdgt_100, y = bdgt_1000)) + 
  labs(x="BdgT - NIVA 100 lakes", y = "BdgT - NIVA 1000 lakes") +
    theme_minimal(base_size = 25)
p3 <- ggplot(data = rr.all) + geom_point(aes(x=TOC_100, y = TOC_1000)) + 
  labs(x="TOC - NIVA 100 lakes", y = "TOC - NIVA 1000 lakes") +
    theme_minimal(base_size = 25)

plot_grid(plotlist = list(p1,p2,p3), nrow = 1, ncol = 3)

```

```{r compare-log, fig.dim = c(18,6)}

p4 <- ggplot(data = rr.all) + geom_point(aes(x=log.rr_100, y = log.rr_1000, col = TOC_100)) + 
  labs(x="log RR - NIVA 100 lakes", y = "log RR - NIVA 1000 lakes") +
    theme_minimal(base_size = 25)

p5 <- ggplot(data = rr.all) + geom_point(aes(x=log.bdgt_100, y = log.bdgt_1000, col = TOC_100)) + 
  labs(x="log BdgT - NIVA 100 lakes", y = "log BdgT - NIVA 1000 lakes") +
    theme_minimal(base_size = 25)

plot_grid(plotlist = list(p4,p5), nrow = 1, ncol = 2)

```

## SCM

from http://www.dagitty.net/dags.html#

```{r DAG-plot}
testImplications <- function( covariance.matrix, sample.size ){
	library(ggm)
	tst <- function(i){ pcor.test( pcor(i,covariance.matrix), length(i)-2, sample.size )$pvalue }
tos <- function(i){ paste(i,collapse=" ") }
implications <- list(c("forest","agriculture"),
		c("forest","peat"),
		c("forest","log.ndep"),
		c("forest","log.runoff"),
		c("forest","log.cond","log.tp","log.tn"),
		c("forest","log.cond","log.toc"),
		c("forest","log.cond","ndvi"),
		c("forest","log.cndom","log.toc"),
		c("forest","log.cndom","ndvi"),
		c("forest","pH","log.toc"),
		c("forest","pH","ndvi"),
		c("forest","log.fe"),
		c("forest","log.soil"),
		c("forest","log.toc","ndvi"),
		c("forest","log.tn","log.toc"),
		c("forest","log.tn","ndvi"),
		c("forest","log.tp","log.toc"),
		c("forest","log.tp","ndvi"),
		c("forest","log.rr","log.cndom"),
		c("forest","log.rr","log.toc"),
		c("forest","log.rr","ndvi"),
		c("forest","log.bdgt","log.cond"),
		c("forest","log.bdgt","log.tp","log.tn"),
		c("forest","log.bdgt","log.toc"),
		c("forest","log.bdgt","ndvi"),
		c("agriculture","peat"),
		c("agriculture","log.ndep"),
		c("agriculture","log.runoff"),
		c("agriculture","log.cond","log.tp","log.tn"),
		c("agriculture","log.cond","log.toc","log.tp"),
		c("agriculture","log.cndom"),
		c("agriculture","pH"),
		c("agriculture","log.fe"),
		c("agriculture","ndvi"),
		c("agriculture","log.soil"),
		c("agriculture","log.toc"),
		c("agriculture","log.tn"),
		c("agriculture","log.rr"),
		c("agriculture","log.bdgt","log.cond"),
		c("agriculture","log.bdgt","log.tp","log.tn"),
		c("agriculture","log.bdgt","log.toc","log.tp"),
		c("peat","log.ndep"),
		c("peat","log.runoff"),
		c("peat","log.cond","log.tp","log.tn"),
		c("peat","log.cond","log.toc"),
		c("peat","log.cndom","log.toc"),
		c("peat","pH","log.toc"),
		c("peat","log.fe"),
		c("peat","ndvi"),
		c("peat","log.soil"),
		c("peat","log.tn","log.toc"),
		c("peat","log.tp","log.toc"),
		c("peat","log.rr","log.cndom"),
		c("peat","log.rr","log.toc"),
		c("peat","log.bdgt","log.cond"),
		c("peat","log.bdgt","log.tp","log.tn"),
		c("peat","log.bdgt","log.toc"),
		c("log.ndep","log.runoff"),
		c("log.ndep","log.cond","log.tp","log.tn"),
		c("log.ndep","log.cond","log.toc","log.tn"),
		c("log.ndep","log.cndom"),
		c("log.ndep","log.fe"),
		c("log.ndep","ndvi"),
		c("log.ndep","log.soil"),
		c("log.ndep","log.toc"),
		c("log.ndep","log.tp"),
		c("log.ndep","log.rr"),
		c("log.ndep","log.bdgt","log.cond"),
		c("log.ndep","log.bdgt","log.tp","log.tn"),
		c("log.ndep","log.bdgt","log.toc","log.tn"),
		c("log.runoff","log.cond"),
		c("log.runoff","log.cndom"),
		c("log.runoff","pH"),
		c("log.runoff","log.fe"),
		c("log.runoff","ndvi"),
		c("log.runoff","log.toc"),
		c("log.runoff","log.tn"),
		c("log.runoff","log.tp"),
		c("log.runoff","log.rr"),
		c("log.runoff","log.bdgt"),
		c("log.cond","log.cndom","log.toc"),
		c("log.cond","log.cndom","log.tp","log.tn"),
		c("log.cond","pH","log.ndep","log.toc"),
		c("log.cond","pH","log.toc","log.tn"),
		c("log.cond","pH","log.tp","log.tn"),
		c("log.cond","log.fe"),
		c("log.cond","ndvi","log.toc"),
		c("log.cond","ndvi","log.tp","log.tn"),
		c("log.cond","log.soil"),
		c("log.cond","log.toc","log.tp","log.tn"),
		c("log.cond","log.rr","log.cndom"),
		c("log.cond","log.rr","log.toc"),
		c("log.cond","log.rr","log.tp","log.tn"),
		c("log.cndom","pH","log.toc"),
		c("log.cndom","ndvi","log.toc"),
		c("log.cndom","log.soil"),
		c("log.cndom","log.tn","log.toc"),
		c("log.cndom","log.tp","log.toc"),
		c("log.cndom","log.bdgt","log.cond"),
		c("log.cndom","log.bdgt","log.tp","log.tn"),
		c("log.cndom","log.bdgt","log.toc"),
		c("pH","log.fe"),
		c("pH","ndvi","log.toc"),
		c("pH","log.soil"),
		c("pH","log.tn","log.toc","log.ndep"),
		c("pH","log.tp","log.toc"),
		c("pH","log.rr","log.cndom"),
		c("pH","log.rr","log.toc"),
		c("pH","log.bdgt","log.cond"),
		c("pH","log.bdgt","log.tp","log.tn"),
		c("pH","log.bdgt","log.toc","log.tn"),
		c("pH","log.bdgt","log.toc","log.ndep"),
		c("log.fe","ndvi"),
		c("log.fe","log.soil"),
		c("log.fe","log.toc"),
		c("log.fe","log.tn"),
		c("log.fe","log.tp"),
		c("log.fe","log.rr","log.cndom"),
		c("log.fe","log.bdgt"),
		c("ndvi","log.soil"),
		c("ndvi","log.tn","log.toc"),
		c("ndvi","log.tp","log.toc"),
		c("ndvi","log.rr","log.cndom"),
		c("ndvi","log.rr","log.toc"),
		c("ndvi","log.bdgt","log.cond"),
		c("ndvi","log.bdgt","log.tp","log.tn"),
		c("ndvi","log.bdgt","log.toc"),
		c("log.soil","log.toc"),
		c("log.soil","log.tn"),
		c("log.soil","log.tp"),
		c("log.soil","log.rr"),
		c("log.soil","log.bdgt"),
		c("log.toc","log.rr","log.cndom"),
		c("log.toc","log.bdgt","log.cond"),
		c("log.toc","log.bdgt","log.tp","log.tn"),
		c("log.tn","log.tp","log.toc"),
		c("log.tn","log.rr","log.cndom"),
		c("log.tn","log.rr","log.toc"),
		c("log.tn","log.bdgt","log.cond"),
		c("log.tp","log.rr","log.cndom"),
		c("log.tp","log.rr","log.toc"),
		c("log.tp","log.bdgt","log.cond"),
		c("log.rr","log.bdgt","log.cond"),
		c("log.rr","log.bdgt","log.tp","log.tn"),
		c("log.rr","log.bdgt","log.toc"),
		c("log.rr","log.bdgt","log.cndom"))
	data.frame( implication=unlist(lapply(implications,tos)),
		pvalue=unlist( lapply( implications, tst ) ) )

}
```

## Transform dataset

Detection and log of skewed data. 

```{r skewness, eval = F}
bdg.data <- readRDS("bdg.data.rds")

bdg.df <- bdg.data[,- which(names(bdg.data) %in% id.names[-which(id.names %in% c("CBA_Lake_ID","longitude","latitude","region","region2"))])] %>% st_drop_geometry()

skewed.data <- find_skewness(bdg.df, index = F)

bdg.trans <- bdg.df
for(i in skewed.data){
  newname <- paste("log",i, sep = ".")
  oldnames <- names(bdg.trans)
  bdg.trans <<- cbind(bdg.trans, log(bdg.trans[[i]])) %>% setNames(c(oldnames,newname))
}

# bdg.trans <- bdg.data %>% mutate_at(.funs = transform, method = "log", .vars = skewed.data) %>% st_drop_geometry()
# sapply(bdg.trans,"class")
# bdg.trans[,which(names(bdg.trans) %in% skewed.data)] <- as.numeric(bdg.trans[,which(names(bdg.trans) %in% skewed.data)])
# 
# skewed.index <- which(names(bdg.trans) %in% skewed.data)
# names(bdg.trans)[skewed.index] <- paste("log",skewed.data,sep = ".")

find_na(bdg.trans, index = F)
bdg.trans <- bdg.trans %>%  filter(is.na(log.ndep) == F) %>% filter(is.na(log.cndom)==F) %>% filter(log.soil != -Inf) %>% filter(log.bdgt != -Inf)

bdg.trans$log.agriculture <- NULL
bdg.trans$log.forest <- NULL
bdg.trans$log.peat <- NULL
bdg.trans$log.ndvi <- NULL

saveRDS(bdg.trans,"bdg.trans.rds")
```

```{r select-variables, eval = T}
bdg.trans <- readRDS("bdg.trans.rds")

vars <- c("region2", "agriculture","forest","peat","log.ndep","log.runoff","log.cond","log.cndom","ph","log.fe","ndvi","log.soil","log.toc","log.tn","log.tp","log.rr","log.bdgt","log.rrn")

bdg.sel <- bdg.trans[vars]
saveRDS(bdg.sel, "bdg.sel.rds")

```

## Explore dataset

## PCA

```{r pca-total, fig.dim = c(10,8)}

library(viridisLite)

pca.rr <- prcomp(formula = ~., data = bdg.sel[,-which(names(bdg.sel) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.12 <- fviz_pca_biplot(pca.rr,
                              habillage = "none",
                             col.ind = bdg.sel$region2, 
                             label = "var",
                             col.var = "black",
                             labelsize = 7,
                             repel = T,arrowsize=1)+theme_bw(base_size=30)

plot.pca.23 <- fviz_pca_biplot(pca.rr, axes = c(2,3),
                              habillage = "none",
                             col.ind = bdg.sel$region2, 
                             label = "var",
                             labelsize = 7,
                             col.var = "black",
                             repel = T,arrowsize=1)+theme_bw(base_size=30)
par(mfrows = c(1,2))
plot.pca.12
plot.pca.23

``` 

### Scatterplot TOC

```{r difference-groups-toc, fig.dim = c(18,4)}
ggplot(bdg.sel, aes(x=log.toc, y = log.bdgt, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size= 25)+theme(legend.position = "none")

ggplot(bdg.sel, aes(x=log.toc, y = log.rr, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size = 25)+theme(legend.position = "none")
```

### Scatterplot Runoff

```{r difference-groups-runoff, fig.dim = c(18,4)}
ggplot(bdg.sel, aes(x=log.runoff, y = log.bdgt, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size= 25)+theme(legend.position = "none")

ggplot(bdg.sel, aes(x=log.runoff, y = log.rr, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size = 25)+theme(legend.position = "none")
```

### Scatterplot Conductivity 

```{r difference-groups-cond, fig.dim = c(18,4)}
ggplot(bdg.sel, aes(x=log.cond, y = log.bdgt, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size= 25)+theme(legend.position = "none")

ggplot(bdg.sel, aes(x=log.cond, y = log.rr, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size = 25)+theme(legend.position = "none")
```

### Scatterplot NDVI

```{r difference-groups-ndvi, fig.dim = c(18,4)}
ggplot(bdg.sel, aes(x=ndvi, y = log.bdgt, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size= 25)+theme(legend.position = "none")

ggplot(bdg.sel, aes(x=ndvi, y = log.rr, col = region2))+geom_point()+geom_smooth(method = "lm")+
  facet_grid(cols = vars(region2))+theme_minimal(base_size = 25)+theme(legend.position = "none")
```

## Corrplot for all lakes

```{r corrplot, eval= T}
bdg.sel <- readRDS("bdg.sel.rds")

cor.mat <- cor(bdg.sel[,-1])
p.mat <- corrplot::cor.mtest(bdg.sel[,-1],conf.level=0.95)
corrplot::corrplot(cor.mat,type=c("lower"), method = "color", addCoef.col = "grey", p.mat = p.mat$p,insig = "blank", diag = F)
```

package
nlmi
lmi
glmmtmb


## SELM

```{r selm-total}

bdg.spdf <- SpatialPointsDataFrame(bdg.trans[,c("longitude","latitude")], bdg.sel)
bdg.kmat <- knearneigh(bdg.spdf, k = 50) %>% knn2nb() %>% nb2listw()

rr.selm <- errorsarlm(formula= log.rr~log.cndom+log.ndep+log.tp, bdg.sel, bdg.kmat)
qplot(rr.selm$fitted.values, bdg.sel$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(y="log.rr for Norway", x = "Predicted values for log.rr") + theme_minimal(base_size = 25)

bdgt.selm <- errorsarlm(formula= log.bdgt~forest + log.runoff, bdg.sel, bdg.kmat)
qplot(bdgt.selm$fitted.values, bdg.sel$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(y="log.bdgt for Norway", x = "Predicted values for log.bdgt") + theme_minimal(base_size = 25)
```

# Analysis by region

```{r dataframes-regions}
bdg.trans <- readRDS("bdg.trans.rds")
bdg.sel <- readRDS("bdg.sel.rds")

b1 <- bdg.sel %>% filter(region2 == "1-South") %>% subset(select = -1)
b1.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "1-South")[,c("longitude","latitude")], b1) 
b1.kmat <- knearneigh(b1.spdf, k = 10) %>% knn2nb() %>% nb2listw()

b2 <- bdg.sel %>% filter(region2 == "2-East") %>% subset(select = -1) 
b2.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "2-East")[,c("longitude","latitude")], b2) 
b2.kmat <- knearneigh(b2.spdf, k = 10) %>% knn2nb() %>% nb2listw()

b3 <- bdg.sel %>% filter(region2 == "3-Mountains") %>% subset(select = -1)
b3.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "3-Mountains")[,c("longitude","latitude")], b3) 
b3.kmat <- knearneigh(b3.spdf, k = 10) %>% knn2nb() %>% nb2listw()

b4 <- bdg.sel %>% filter(region2 == "4-West") %>% subset(select = -1)
b4.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "4-West")[,c("longitude","latitude")], b4) 
b4.kmat <- knearneigh(b4.spdf, k = 10) %>% knn2nb() %>% nb2listw()

b5 <- bdg.sel %>% filter(region2 == "5-Central") %>% subset(select = -1)
b5.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "5-Central")[,c("longitude","latitude")], b5) 
b5.kmat <- knearneigh(b5.spdf, k = 10) %>% knn2nb() %>% nb2listw()

b6 <- bdg.sel %>% filter(region2 == "6-North") %>% subset(select = -1)
b6.spdf <- SpatialPointsDataFrame(filter(bdg.trans, region2 == "6-North")[,c("longitude","latitude")], b6) 
b6.kmat <- knearneigh(b6.spdf, k = 10) %>% knn2nb() %>% nb2listw()
```

## ANOVA

```{r boxplot-by-region, fig.dim = c(10,5)}

#ggplot(bdg.trans,aes(x = region2, y=log.rrn))+geom_boxplot(outlier.shape = NA, notch = T)+
#  geom_jitter(aes(col=region2))+ labs(x = "", y = "log.rrn", col = "Region")+
#  theme_minimal(base_size = 25)

ggplot(bdg.trans,aes(x = region2, y=log.rr))+geom_boxplot(outlier.shape = NA, notch = T)+
  geom_jitter(aes(col=region2))+ labs(x = "", y = "log.rr", col = "Region")+
  theme_minimal(base_size = 25)

ggplot(bdg.trans,aes(x = region2, y=log.bdgt))+geom_boxplot(outlier.shape = NA, notch = T)+
  geom_jitter(aes(col=region2))+ labs(x = "", y = "log.bdgt", col = "Region")+
  theme_minimal(base_size = 25)
```
```{r anova}

rr.aov <- aov(log.rr~region2, data = bdg.sel) 

summary(rr.aov)[[1]] %>% as.data.frame() %>% knitr::kable(caption = "ANOVA log.rr") %>% 
  kable_styling(bootstrap_options = "bordered", position = "center", full_width = F)
rr.hsd <- TukeyHSD(rr.aov)

rr.hsd.df <- rr.hsd$region2 %>% as.data.frame %>% setNames(c("diff","lwr","upr","p.adj"))

rr.hsd.df %>%
  knitr::kable(format = "html", digits = 3, caption = "Tukey HSD for log.rr ~ region") %>% 
  row_spec(which(rr.hsd.df$p.adj < 0.05), color = "red") %>%
  kable_styling(bootstrap_options = "bordered", position = "center", full_width = F)


bdgt.aov <- aov(log.bdgt~region2, data = bdg.sel)
summary(bdgt.aov)[[1]] %>% as.data.frame() %>% knitr::kable(caption = "ANOVA log.bdgt") %>% 
  kable_styling(bootstrap_options = "bordered", position = "center", full_width = F)
bdgt.hsd <- TukeyHSD(bdgt.aov)

bdgt.hsd.df <- bdgt.hsd$region2 %>% as.data.frame %>% setNames(c("diff","lwr","upr","p.adj"))

bdgt.hsd.df %>%
  knitr::kable(format = "html", digits = 3, caption = "Tukey HSD for log.bdgt ~ region") %>% 
  row_spec(which(bdgt.hsd.df$p.adj < 0.05), color = "red") %>%
  kable_styling(bootstrap_options = "bordered", position = "center", full_width = F)

```

## PCA 

```{r pca-region, fig.dim = c(24,18)}
pca.b1 <- prcomp(formula = ~., data = b1[,-which(names(b1) %in% c("region2", "log.rrn"))], center = T, scale. = T)

library(scales)
col.pca.vars <- hue_pal()(6)

plot.pca.b1 <- fviz_pca_biplot(pca.b1, title = "PCA 1-South",
                              habillage = "none", label = "var", col.var = col.pca.vars[1], 
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)+theme(legend.position = "none")
    

pca.b2 <- prcomp(formula = ~., data = b2[,-which(names(b2) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.b2 <- fviz_pca_biplot(pca.b2, title = "PCA 2-East",
                              habillage = "none", label = "var", col.var = col.pca.vars[2],
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)

pca.b3 <- prcomp(formula = ~., data = b3[,-which(names(b3) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.b3 <- fviz_pca_biplot(pca.b3, title = "PCA 3-Mountains",
                              habillage = "none", label = "var",col.var = col.pca.vars[3],
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)
  
pca.b4 <- prcomp(formula = ~., data = b4[,-which(names(b4) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.b4 <- fviz_pca_biplot(pca.b4,title = "PCA 4-West",
                              habillage = "none", label = "var", col.var = col.pca.vars[4],
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)

pca.b5 <- prcomp(formula = ~., data = b5[,-which(names(b5) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.b5 <- fviz_pca_biplot(pca.b5, title = "PCA 5-Central",
                              habillage = "none", label = "var", col.var = col.pca.vars[5],
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)

pca.b6 <- prcomp(formula = ~., data = b6[,-which(names(b6) %in% c("region2", "log.rrn"))], center = T, scale. = T)

plot.pca.b6 <- fviz_pca_biplot(pca.b6, title = "PCA 6-North",
                              habillage = "none", label = "var", col.var = col.pca.vars[6],
                             repel = T,arrowsize=1,labelsize=10)+theme_bw(base_size = 35)


plot_grid(plotlist = list(plot.pca.b1, plot.pca.b2, plot.pca.b3, plot.pca.b4, plot.pca.b5, plot.pca.b6), nrow = 2, ncol = 3)
```

## Corrplots

```{r corrplot-by-region, fig.dim = c(24,20)}

par(mfrow = c(2,3), cex = 2.5)

p.mat1 <- corrplot::cor.mtest(b1,conf.level=0.95)
corrplot(cor(b1), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat1$p, insig = "blank", diag = F, title = "Region 1: South", mar = c(0,0,1,0))

p.mat2 <- corrplot::cor.mtest(b2,conf.level=0.95)
corrplot(cor(b2), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat2$p, insig = "blank", diag = F, title = "Region 2: East", mar = c(0,0,1,0))

p.mat3 <- corrplot::cor.mtest(b3,conf.level = 0.95)
corrplot(cor(b3), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat3$p, insig = "blank", diag = F, title = "Region 3: Mountains", mar = c(0,0,1,0))

p.mat4 <- cor.mtest(b4,conf.level = 0.95)
corrplot(cor(b4), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat4$p, insig = "blank", diag = F, title = "Region 4: West", mar = c(0,0,1,0))

p.mat5 <- cor.mtest(b5, conf.level = 0.95)
corrplot(cor(b5), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat5$p, insig = "blank", diag = F, title = "Region 5: Central", mar = c(0,0,1,0))

p.mat6 <- cor.mtest(b6, conf.level = 0.95)
corrplot(cor(b6), type = "lower", method = "color", addCoef.col = "grey", p.mat = p.mat6$p, insig = "blank", diag = F, title = "Region 6: North", mar = c(0,0,1,0))
```

```{r table-resuts}
data.frame(region = c("Global","South", "East","Moutains","West","Central","North"), 
           cor.rr = c("forest, log.ndep, log.cndom, ph, log.fe, log.toc, log.tn, log.tp",
                      "log.fe, log.toc, log.tn", 
                      "log.fe, log.toc, log.tn",
                      "NULL",
                      "log.cndom, log.toc",
                      "peat, log.cond, log.fe, log.soil, log.toc, log.tn, log.tp", 
                      "log.ndep, log.runoff, log.cndom"), 
           cor.bdgt = c("forest, log.ndep, log.runoff",
                        "forest, log.ndep, log.runoff, ndvi, log.soil",
                        "forest, log.ndep, log.runoff, ndvi, log.soil",
                        "NULL",
                        "log.cond,log.cndom,ph,log.fe,log.toc,log.tn",
                        "NULL",
                        "log.ndep, log.runoff, log.cond, ph")) %>% 
  kable %>% kable_styling(bootstrap_options = "bordered", position = "center", full_width = F)

```

## Scatterplots

```{r scatterplots, fig.dim = c(9,12)}
par(mfrow = c(4,3), cex.lab = 3, cex.main = 3)
plot(log.rr~forest+log.ndep+log.cndom+ph+log.fe+log.toc+log.tn+log.tp, data = bdg.sel, main = "Global, log.rr", col = "cornflowerblue")
plot(log.bdgt~forest+log.ndep+log.runoff, data = bdg.sel, main = "Global, log.bdgt", col = "brown3")

```

```{r scatterplots-b1, fig.dim = c(9,9)}
par(mfrow = c(3,3), cex.lab = 3, cex.main = 3)
plot(log.rr~log.fe + log.toc+log.tn, data = b1, main = "1- South, log.rr", col = "cornflowerblue")
plot(log.bdgt~forest + log.ndep + log.runoff + ndvi + log.soil, data = b1, main = "1- South, log.bdgt", col = "brown3")
```

```{r scatterplots-b2, fig.dim = c(9,9)}
par(mfrow = c(3,3), cex.lab = 3, cex.main = 3)
plot(log.rr~log.fe + log.toc+log.tn, data = b2, main = "2- East, log.rr", col = "cornflowerblue")
plot(log.bdgt~forest+log.ndep+log.runoff+ndvi+log.soil, data = b2, main = "2- East, log.bdgt", col = "brown3")
```

```{r scatterplots-b3, fig.dim = c(9,3)}
par(mfrow = c(1,3), cex.lab = 3, cex.main = 3)
plot(log.bdgt~log.soil, data = b3, main = "3- Mountains, log.bdgt", col = "brown3")
```

```{r scatterplots-b4, fig.dim = c(9,9)}
par(mfrow = c(3,3), cex.lab = 3, cex.main = 3)
plot(log.rr~log.cndom + log.toc, data = b4, main = "4- West, log.rr", col = "cornflowerblue")
plot(log.bdgt~log.cndom + ph + log.fe + log.toc + log.tn, data = b2, main = "4- West, log.bdgt", col = "brown3")
```

```{r scatterplots-b5, fig.dim = c(9,6)}
par(mfrow = c(2,3), cex.lab = 3, cex.main = 3)
plot(log.rr~peat + log.fe + log.soil + log.toc + log.tn + log.tp, data = b5, main = "5- Central, log.rr", col = "cornflowerblue")
```

```{r scatterplots-b6, fig.dim = c(9,9)}
par(mfrow = c(3,3), cex.lab = 3, cex.main = 3)
plot(log.rr~ log.ndep + log.runoff + log.cndom, data = b6, main = "6- North, log.rr", col = "cornflowerblue")
plot(log.bdgt~ log.ndep + log.runoff + log.cond + ph, data = b6, main = "6- North, log.bdgt", col = "brown3")
```

## SELM Regions

Formula: ~ log.toc + log.cond + ndvi + log.runoff 

```{r selm-regions, fig.dim = c(15,30)}

rr.fm <- "log.rr~log.toc + log.cond + ndvi + log.runoff" %>% as.formula()
bdgt.fm <- "log.bdgt~log.toc + log.cond + ndvi + log.runoff" %>% as.formula()

b1.rr.selm <- errorsarlm(formula= rr.fm, b1, b1.kmat)
p1.rr <- qplot(b1.rr.selm$fitted.values, b1$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y="log.rr", x = "Predicted log.rr", title = "South - SELM log.rr")+
  theme_minimal(base_size = 30)

b1.bdgt.selm <- errorsarlm(formula= bdgt.fm, b1, b1.kmat)
p1.bdgt <- qplot(b1.bdgt.selm$fitted.values, b1$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y="log.bdgt", x = "Predicted log.bdgt", title = "South - SELM log.bdgt") +
  theme_minimal(base_size = 30)

###
b2.rr.selm <- errorsarlm(formula= rr.fm, b2, b2.kmat)
p2.rr <- qplot(b2.rr.selm$fitted.values, b2$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y ="log.rr", x = "Predicted log.rr", title = "East - SELM log.rr")+
  theme_minimal(base_size = 30)

b2.bdgt.selm <- errorsarlm(formula= bdgt.fm, b2, b2.kmat)
p2.bdgt <- qplot(b2.bdgt.selm$fitted.values, b2$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") + labs(y="log.bdgt", x = "Predicted log.bdgt", title = "East - SELM log.bdgt")+
  theme_minimal(base_size = 30)

###
b3.rr.selm <- errorsarlm(formula= rr.fm, b3, b3.kmat)
p3.rr <- qplot(b3.rr.selm$fitted.values, b3$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y="log.rr", x = "Predicted log.rr", title = "Moutains - SELM log.rr")+
  theme_minimal(base_size = 30)

b3.bdgt.selm <- errorsarlm(formula= bdgt.fm, b3, b3.kmat)
p3.bdgt <- qplot(b3.bdgt.selm$fitted.values, b3$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") + labs(y="log.bdgt", x = "Predicted log.bdgt", title = "Mountains - SELM log.bdgt")+
  theme_minimal(base_size = 30)

###
b4.rr.selm <- errorsarlm(formula= rr.fm, b4, b4.kmat)
p4.rr <- qplot(b4.rr.selm$fitted.values, b4$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y ="log.rr", x = "Predicted log.rr", title = "West - SELM log.rr")+
  theme_minimal(base_size = 30)

b4.bdgt.selm <- errorsarlm(formula= bdgt.fm, b4, b4.kmat)
p4.bdgt <- qplot(b4.bdgt.selm$fitted.values, b4$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") + labs(y="log.bdgt", x = "Predicted log.bdgt", title = "West - SELM log.bdgt")+
  theme_minimal(base_size = 30)

###
b5.rr.selm <- errorsarlm(formula= rr.fm, b5, b5.kmat)
p5.rr <- qplot(b5.rr.selm$fitted.values, b5$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y="log.rr", x = "Predicted log.rr", title = "Central - SELM log.rr")+
  theme_minimal(base_size = 30)

b5.bdgt.selm <- errorsarlm(formula= bdgt.fm, b5, b5.kmat)
p5.bdgt <- qplot(b5.bdgt.selm$fitted.values, b5$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") + labs(y ="log.bdgt", x = "Predicted log.bdgt", title = "Central - SELM log.bdgt")+
  theme_minimal(base_size = 30)

###
b6.rr.selm <- errorsarlm(formula= rr.fm, b6, b6.kmat)
p6.rr <- qplot(b6.rr.selm$fitted.values, b6$log.rr)+geom_abline(slope = 1, intercept = 0, col = "red")+labs(y="log.rr", x = "Predicted log.rr", title = "North - SELM log.rr")+
  theme_minimal(base_size = 30)

b6.bdgt.selm <- errorsarlm(formula= bdgt.fm, b6, b6.kmat)
p6.bdgt <- qplot(b6.bdgt.selm$fitted.values, b6$log.bdgt)+geom_abline(slope = 1, intercept = 0, col = "red") + labs(y="log.bdgt", x = "Predicted log.bdgt", title = "North - SELM log.bdgt")+
  theme_minimal(base_size = 30)

plot_grid(plotlist = list(p1.rr, p1.bdgt, p2.rr, p2.bdgt, p3.rr, p3.bdgt, p4.rr, p4.bdgt, p5.rr, p5.bdgt, p6.rr, p6.bdgt), nrow = 6, ncol = 2)
```





`r knitr::knit_exit()`

```{r lm by region}
rr.lm <- lm(log.rr ~ forest + log.ndep + log.cndom + log.soil, data = bdg.sel ) # ph cor with log.ndep, fe+toc+tn cor with cndom
rr.lm1 <- lm(log.rr ~ log.toc, data = b1) # fe, toc and tn are correlated ~0.95
rr.lm2 <- lm(log.rr ~ log.toc, data = b2)
rr.lm3 <- lm(log.rr ~ log.toc, data = b3)
rr.lm4 <- lm(log.rr ~ log.cndom, data = b4)
rr.lm5 <- lm(log.rr ~ peat + log.toc, data = b5)
rr.lm6 <- lm(log.rr ~ log.runoff + log.cndom, data = b6)

bdgt.lm <- lm(log.bdgt ~ forest + log.ndep + log.runoff + log.soil, data = bdg.sel ) # ph cor with log.ndep, fe+toc+tn cor with cndom
bdgt.lm1 <- lm(log.bdgt ~ forest + log.ndep + log.runoff, data = b1) # fe, toc and tn are correlated ~0.95
bdgt.lm2 <- lm(log.bdgt ~ forest + log.ndep + log.runoff, data = b2)
bdgt.lm3 <- lm(log.bdgt ~ ., data = b3)
bdgt.lm4 <- lm(log.bdgt ~ log.cndom + ph + log.tn, data = b4)
bdgt.lm5 <- lm(log.bdgt ~ ., data = b5)
bdgt.lm6 <- lm(log.bdgt ~ log.runoff, log.cond, data = b6)


```



## Model by region

Based on parameters from articles 1 and 2. Catchment parameters are always: tn, peat, runoff, ndvi, agriculture.

Chemistry parameters are 
For RR: log(C/N), logDP and SUVA
For RRn: SUVA, log(Fe), SARuv, logDP
For BdgT: logDOC, pH, SARuv, bacterial abundance. 

We don't have absorbance data so SARuv and SUVA are replaced by cndom.

South (region 4+5), East (1+2), Inland (3), West (region 6+7), Central (region 8) and North (region 9+10)

```{r region-1}

bdg1 <- filter(bdg.sel,region == 1)
plot(formula = log.rr ~. , data = bdg1)

bdg2 <- filter(bdg.sel,region == 2)
plot(formula = log.rr ~. , data = bdg2)

bdg3 <- filter(bdg.sel,region == 3)
plot(formula = log.rr ~. , data = bdg3)

bdg4 <- filter(bdg.sel,region == 4)
plot(formula = log.rr ~. , data = bdg4)

bdg5 <- filter(bdg.sel,region == 5)
plot(formula = log.rr ~. , data = bdg5)

bdg6 <- filter(bdg.sel,region == 6)
plot(formula = log.rr ~. , data = bdg6)

bdg7 <- filter(bdg.sel,region == 7)
plot(formula = log.rr ~. , data = bdg7)

bdg8 <- filter(bdg.sel,region == 8)
plot(formula = log.rr ~. , data = bdg8)

bdg9 <- filter(bdg.sel,region == 9)
plot(formula = log.rr ~. , data = bdg9)

bdg10 <- filter(bdg.sel,region == 10)
plot(formula = log.rr ~. , data = bdg10)

bdg.south <- filter(bdg.sel, region %in% c(4,5))
plot(formula = log.rr~., data = bdg.south)

bdg.east <- filter(bdg.sel, region %in% c(1,2))
plot(formula = log.rr~., data = bdg.east)

bdg.inland <- filter(bdg.sel, region %in% c(3))
plot(formula = log.rr~., data = bdg.inland)

bdg.west <- filter(bdg.sel, region %in% c(6,7))
plot(formula = log.rr~., data = bdg.west)

bdg.central <- filter(bdg.sel, region %in% c(8))
plot(formula = log.rr~., data = bdg.central)

bdg.north <- filter(bdg.sel, region %in% c(9,10))
plot(formula = log.rr~., data = bdg.central)

lm(formula = log.rr ~ ndvi + log.ndep + peat + agriculture + runoff + cndom + log.toc + log.tn + log.tp , data = bdg1) %>% summary()
lm(formula = log.rr ~ peat + cndom, data = bdg1) %>% summary()

gam(formula = log.rr ~ ndvi + log.ndep + peat + agriculture + runoff + cndom + log.toc + log.tn + log.tp , data = bdg1) %>% summary()

full.model <- lm(log.rr~ndvi + log.ndep + peat + agriculture + runoff + cndom + log(toc/tn) + log.tp, data = bdg1) 
summary(full.model)
step.model <- stepAIC(full.model, direction = "both")
summary(step.model)

final.model <- lm(log.rr~peat+cndom, data = bdg1)
summary(final.model)


```

## 